### 2014-08-05 paper/paper.tex
Made Meg's edits. Up to 18 pages

%\documentclass[12pt]{article}
\documentclass[final,3p,12pt]{elsarticle}
\usepackage{enumerate}
\usepackage{courier}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{float}
\usepackage{setspace}
\usepackage{fancyvrb}
\usepackage[numbers]{natbib}

@@

Draft 3 %TODO ditch this

@@

Social media allows people to discuss topics and disseminate their opinions
with a near negligible cost. The low barrier of entry to social media allows
everyone to share their opinions as much as they please. The scale of textual
data from social media is enormous. In order to gain insight into the `overall'
opinion on particular topics, the process of determining sentiment must be
automated.

@@

Sentiment analysis is simply a classification task that attempts to label
fragments of text that appropriately summarizes the attitude of the text.
Natural Language Processing (NLP) provides a number of method for preforming
sentiment analysis. Many of these methods, such as Naive Bayes and Maximum
Entropy, take a probabilistic approach to performing the task. Another commonly
used method of performing sentiment analysis is the uses of a support vector
machine (SVM).

@@

The focus of this undergraduate research project has been to preform sentiment
analysis in data from the microbloging website Twitter. All of the messages,
known as Tweets, are constrained in length to 140 characters. This limit forces
many users of Twitter to use abbreviations and contractions. This data presents
some interesting problems to performing good sentiment analysis due to it's
short and informal nature. Twitter makes it extremely easy to publish a Tweet
and has no way to edit a Tweet once published. This results in more spelling
mistakes and inconsistent grammar then other sources of text. Tweets often also
contain URLs to other websites. Tweets also contain a certain amount of OOV
(Out-Of-Vocabulary) words, such as Hashtags, a tagging system for topics
allowing Tweets in a similar vein of conversation to be found. Other OOV words
include mentions (e.g. `@BarackObama') which are a method to direct a Tweet to
one or more users, as well as `RT' which is an acronym for `retweet'
communicating that this particular Tweet is a reissuing of another users Tweet.
Another issue that is out of the scope of this project is that Twitter
currently supports 35+ languages, meaning that in-order to determine the
sentiment most systems must first determine which language the Tweet is in.

@@

Organizers of the SemEval-2013 workshop created various sentiment analysis
tasks to be preformed. The task focused on for this project was task-2b:
Message polarity classification. The organizers used the Amazon service
``Mechanical Turk" to get many humans to annotate English language Tweets as
`positive', `negative' or `neutral'. In Tweets with mixed sentiments, the
prevailing sentiment was chosen. Participants were tasked with creating a
classifier to establish labels for Tweets not in the training dataset. The
various submissions of the participants were then evaluated by the average of
the F-positive and the F-negative. The submissions were allowed categorized as
`unconstrained' or `constrained' depending on if they did or did not use
supplemental data, respectively.

@@

In the case of sentiment analysis the class labels being applied to the records
denote the attitude or opinion in the text. The focus of this project is
determining sentiment at the document level which assumes that the document
expresses a predominate sentiment. This is a somewhat reasonable assumption
given the fact that Tweets are limited in size, making it difficult to express
mixed sentiment. Sentiment analysis is very closely linked with opinion mining,
which is the practice of extracting sentiment on a particular topic in order to
gain insights into public opinion on the topic.

Much of the previous research make use of Naive Bayes, Maximum Entropy, and
SVM classifiers. The Naive Bayes classifier uses the occurrences of features in
the \textit{training set} to calculate the probability that an example in the
\textit{test set} belongs in a given class. This method has the benefit of
linear time training since there is no iterative process, but I makes the
assumption that all features are independent, which is not the case in natural
language. The Maximum Entropy approach is explained in detail in section
\ref{ssection:summaryofmaxent}. While Maximum Entropy and Naive Bayes are both
probabilistic approaches the to problem, Support Vector Machine (SVM)
classifiers take a different approach to the problem. ``SVM Classifiers attempt
partition the data space with the use of linear or non-linear delineations
between the different classifiers.''\cite{Aggarwal2012}.

In order to perform text classification of any kind one of the most important
aspects is deciding how to represent a document and how to select
representative features. A document is often represented as either a
\textit{bag-of-words} or \textit{string}. For the limited time frame of this
project the \textit{bag-of-words} method was used.

There are many established methods of selecting features such as document
frequency, which is simply the number of documents that a word occurs at least
once in. Document frequency has the added benefit of not requiring a class
labels to select features. While document frequency is the only method employed
for this project many other method exist, such as Gini Index, Information Gain,
and $\chi^{2}$-statistic \cite{Aggarwal2012}. Each of these more complex
methods rely on the class label in the training set in order to determine how
well the feature discriminates between classes.

@@

The Maximum Entropy classification method was focused on due to the limited
duration of the project, however a Naive Bayes classifier was also implemented
to provide a point of comparison. This supervised machine learning method
calculates the probability that a given document $d$ belongs to the class $c$.
The Maximum Entropy approach is summarized below. For a more detailed
description, see: Nigam et al. \cite{Nigam1999}.

@@

            \item \textbf{Numbers} are replaced with "\#NUM"

            \item \textbf{User} are replaced with "@USER"

            \item \textbf{Hashtags} are replaced with "\#OCTOTHORPE"

            \item \textbf{URLs} are replaced with "@URL"

            \item \textbf{Emoticon} are grouped into four different groups
                based on their intended meaning. Each group is then replaced
                with a representation. (e.g. `:)' $\rightarrow$ `EM\_HAPPY' and
                `;-p' $\rightarrow$ `EM\_WINK')

            \item most \textbf{Punctuation} is replaced with "PUNCT" however
                question marks, punctuation marks, and ellipsis are each
                replaced with their own representation.

@@

            \item By \textbf{Removing Stopwords}, words that do not
                discriminate between classes are eliminated. Because of the
                limited size of the dataset only 28 stopwords are used. For a
                full list of stopwords see section \ref{ssection:stopwords}

@@

The implementation created for this project used a \textit{validation set} during
training as the primary mechanism to avoid overfitting. While training the
effectiveness of the classifier was evaluated after every iteration, using the
accuracy or F-positive measure. During the training process the following
procedure was followed:

@@

        \item \textbf{Folding} - The training and testing process are run
            multiple times with sections used for various roles. The roles are
            permuted so that the classifier is trained and tested with the
            sections in all possible configurations. In each fold 60\% of the
            partitions are designated as part of the \textit{training set},
            20\% as the \textit{validation set}, and 20\% as the \textit{test set}.
            Since the Naive Bayes classifier does not use a validation set
            40\% of the partitions are designated as part of the \textit{test set}

@@

The baseline for this project is established by running the Maximum Entropy
classifier without any stopword removal, uncommon word removal (words with a
document frequancy > 1), Or feature normalization. The F-score that is achieved
is \textbf{50.72}.

@@

\subsubsection{Number of features}
\label{ssection:numfeatures}

Number of features retained after feature selection varies from Tweet to Tweet.
To summarize the number of features retained after different process Table
\ref{table:numfeatures} presents the mean number of features per Tweet for
positive, neutral, negative examples as well as an overall average.

@@

As presented in Table \ref{table:numfeatures}, the number of features decreases
each time a feature set simplification method is turned on. Normalization does
not appear to significantly reduce the number of features per Tweet. This make
sense because the number of features in would only be reduced if it contained
tokens that were different before simplification and were the same after (e.g.
A Tweet that contained `2' and `3' before normalization would only contain the
feature `\#NUM' after normalization).

%TODO clarify

@@

As Table \ref{table:maxent-results} shows `All' experiment provides the best
results over the baseline. The `Just Stopwords' experiment actually provides
slightly worse results then the baseline indicating that in context of Tweets
with so few features in each example every feature counts. The `Just Uncommon'
experiment does not provide very significant gains when used alone. This is
potentially due to the fact that with out normalization turned, features such
as `just' are `Just' are determined to be different and may be removed if each
only occur once.

@@

@@

The implementation create for this project has the ability to show which
features are the most discriminatory. The following is a list of the top 30
most informative features from one run of the Maximum Entropy classifier:

@@

Some of the features listed above make perfect sense, such as `excited' not
being associated with the neutral class and `shitty' being associated with the
negative class. However some of the most discriminating features appear in the
dataset just enough to get past the uncommon word cutoff, such as `calvin' and
`oneil'.

@@

Table \ref{table:comparison} shows the performance of the classifier produced
for this project compared to results of the workshop participants. The full
table (with out this project's results) can be found in table 9 of Nakov et al.
\cite{Nakov2013}.

@@

As shown in Table \ref{table:comparison} there is a lot of room for improvement
before this system is a completive solution. However, only simple methods have
been employed including: stopword removal of only 28 words, words with a
document frequency lower than one removed, and normalization of words, emoticons,
and OOV words. These simple preprocessing and feature selection methods in
combination with a Maximum Entropy classifier have demonstrated reasonable
results.

@@

@@

While a simple approach could be to strip the octothorpe from the front of the
Hashtag (e.g. `\#good' $\rightarrow$ `good'), the semantic meaning of the two
are different. Hashtags often have message level meaning while simple words
have local context withing the sentence. Since the dataset created for the
SemEval-2013 workshop mostly contained examples of Hashtags used as proper
nouns, more annotated Tweets would be needed for training and testing.

As well Hashtags sometimes use CamelCase to create multi-word tags (e.g.
`\#SorryNotSorry'). It could prove worthwhile to attempt the decompose
CamelCase Hashtags into its component words during feature selection.

The top solution from the National Research Council of Canada made use of a
Hashtag lexicon that they developed between April of 2012 and December of the
same year \cite{Mohammad2013}. The final \textit{NRC Hashtag Sentiment Lexicon}
has more than 350,00 entries and is available at
\href{http://www.saifmohammad.com/WebDocs/NRC-Hashtag-Sentiment-Lexicon-v0.1.zip}{Saif
Mohammad's website}.

@@

A challenge that this method could introduce is the introduction of a temporal
element since people's outlooks, and therefore sentiment's in Tweets, will
change over time. As well as public perception of a individual changes, the
prevailing sentiment my change as well.

@@

One problem with the current implementation is that it lack support for
negation. In it's current form the classifier would likely misunderstand the
sequence ``not perfect'' since `not' is a vaguely negative word and `perfect'
is a very positive word, it would likely be classified as positive when in fact
the sentiment is quite clearly negative. A common approach to this problem is
to add a suffix (such as "\_NEG") to word that appear between a negation word
and a section delimiting punctuation mark (such as `,', `.', `:', `;', `!',
`?'). By separating negated features from their non-negated instances, features
can do a better job of distinguishing classes.

@@

One way that many of the other bag-of-words solutions capture word order is by
creating features consisting of sequences of words called bigrams. Adding
bigram features adds a large number of features which can add complexity to
training. This concept can also be extended to sequences of length n, called
ngrams.

@@

Most of the more complex solutions submitted to the SemEval-2013 workshop made
uses of sentiment lexicons that capture a term's polarity. These lexicons, such
as \href{http://sentiwordnet.isti.cnr.it/}{SentiWordNet}, express the positive,
neutral, and negative sentiment of each homonym.

@@

Many of the top solutions submitted to the workshop such as NRC-Canada
\cite{Mohammad2013} made use of a support vector machine for classification.
This indicates that when used effectively a SVM solution can out-perform a
Maximum Entropy solution. The KLUE team on the other hand found that their
Maximum Entropy solution outperform their linear SVM solution
\cite{Proisl2013}.

@@

The list of stopwords was acquired \href{www.ranks.nl/stopwords}{here} and is
as follows:

a, about, an, are, as, at, be, by, com, for, from, how, in, is, it, of, on, or,
that, the, this, to, was, what, when, where, who, will, with

>>>>

%\documentclass[12pt]{article}
\documentclass[final,3p,12pt]{elsarticle}
\usepackage{enumerate}
\usepackage{courier}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{float}
\usepackage{setspace}
\usepackage{fancyvrb}
\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage[numbers]{natbib}

@@

Draft 4 %TODO ditch this

@@

Social media allows people to discuss topics and disseminate their opinions
with a near negligible cost. Social media's low barrier of entry allows
everyone to share their opinions as much as they please. The scale of textual
data from social media is enormous. In order to gain insight into the `overall'
opinion on particular topics, the process of determining sentiment must be
automated.

@@

Sentiment analysis is simply a classification task that attempts to correctly
label fragments of text to appropriately summarizes the attitude of the text.
Natural Language Processing (NLP) provides a number of method for performing
sentiment analysis. Many of these methods, such as Naive Bayes and Maximum
Entropy, take a probabilistic approach to performing the task. Another commonly
used method of performing sentiment analysis is the uses of a support vector
machine (SVM).

@@

The focus of this undergraduate research project has been to perform sentiment
analysis in data from the microbloging website Twitter. All of the messages,
known as Tweets, are constrained in length to 140 characters. This limit forces
many users of Twitter to use abbreviations and contractions. This data presents
some interesting problems to performing good sentiment analysis due to the
short and informal nature of the text. Twitter makes it extremely easy to
publish a Tweet and has no way to edit a Tweet once published. This results in
more spelling mistakes and inconsistent grammar then other sources of text.
Tweets often also contain URLs to other websites. Tweets also contain a certain
amount of OOV (Out-Of-Vocabulary) words, such as Hashtags, a tagging system for
topics allowing Tweets in a similar vein of conversation to be found. Other OOV
words include mentions (e.g. `@BarackObama') which are a method to direct a
Tweet to one or more users, as well as `RT' which is an acronym for `retweet'
communicating that this particular Tweet is a reissuing of another users Tweet.
Another issue that is out of the scope of this project is that Twitter
currently supports 35+ languages, meaning that in order to determine the
sentiment most systems must first determine which language the Tweet is in.

@@

Organizers of the SemEval-2013 workshop created various sentiment analysis
tasks to be preformed. The task focused on for this project was task-2b:
Message polarity classification. The organizers used the Amazon service
``Mechanical Turk" to get many humans to annotate English language Tweets as
`positive', `negative' or `neutral'. In Tweets with mixed sentiments, the
prevailing sentiment was chosen. Participants were tasked with creating a
classifier to establish labels for Tweets not in the training dataset. The
various submissions of the participants were then evaluated by the average of
the F-positive and the F-negative (explained in depth in Section
\ref{ssection:performancemeasurement}). The submissions were allowed
categorized as `unconstrained' or `constrained' depending on if they did or did
not use supplemental data, respectively.

@@

In the case of sentiment analysis, the class labels being applied to the
records denote the attitude or opinion in the text. The focus of this project
is determining sentiment at the document level, which assumes that the document
expresses a predominate sentiment. This is a somewhat reasonable assumption
given the fact that Tweets are limited in size, making it difficult to express
mixed sentiment. Sentiment analysis is very closely linked with opinion mining,
which is the practice of extracting sentiment on a particular topic in order to
gain insights into public opinion on the topic.

Much of the previous research make use of Naive Bayes, Maximum Entropy, and SVM
classifiers. The Naive Bayes classifier uses the occurrences of features in the
\textit{training set} to calculate the probability that an example in the
\textit{test set} belongs in a given class. This method has the benefit of
linear time training since there is no iterative process, but it makes the
assumption that all features are independent, which is not the case in natural
language. The Maximum Entropy approach is explained in detail in section
\ref{ssection:summaryofmaxent}. While Maximum Entropy and Naive Bayes are both
probabilistic approaches the to problem, Support Vector Machine (SVM)
classifiers take a different approach to the problem. ``SVM Classifiers attempt
partition the data space with the use of linear or non-linear delineations
between the different classifiers.''\cite{Aggarwal2012}.

In order to perform text classification of any kind one of the most important
aspects is deciding how to represent a document and how to select
representative features. A document is often represented as a
\textit{bag-of-words}, which is defined a set of strings with no apparent
order. The \textit{bag-of-words} method of representation was selected for this
project.

There are many established methods of selecting features such as document
frequency, which is simply the number of documents that a word occurs at least
once in. Document frequency has the added benefit of not requiring a class
labels to select features. While document frequency is the only method employed
for this project, many other methods exist, such as Gini Index, Information Gain,
and $\chi^{2}$-statistic \cite{Aggarwal2012}. Each of these more complex
methods rely on the class label in the training set in order to determine how
well the feature discriminates between classes.

@@

The Maximum Entropy classification method was focused on due to the limited
duration of the project, however a Naive Bayes classifier was also implemented
to provide a point of comparison. This supervised machine learning method
calculates the probability that a given document $d$ belongs to the class $c$.
The Maximum Entropy approach is summarized below (for a more detailed
description, see: Nigam et al. \cite{Nigam1999}).

@@

            \item \textbf{Numbers} are replaced with ``\#NUM''

            \item \textbf{User} are replaced with ``@USER''

            \item \textbf{Hashtags} are replaced with ``\#OCTOTHORPE''

            \item \textbf{URLs} are replaced with ``@URL''

            \item \textbf{Emoticon} are grouped into four different groups
                based on their intended meaning. Each group is then replaced
                with a representation. [e.g. `:)' $\rightarrow$ `EM\_HAPPY' and
                `;-p' $\rightarrow$ `EM\_WINK']

            \item most \textbf{Punctuation} is replaced with ``PUNCT'' however
                question marks, punctuation marks, and ellipsis are each
                replaced with their own representation.

@@

            \item By \textbf{Removing Stopwords}, words that do not
                discriminate between classes are eliminated. Because of the
                limited size of the dataset only 28 stopwords are used. For a
                full list of stopwords see section \ref{ssection:stopwords}.

@@

The implementation created for this project used a \textit{validation set} during
training as the primary mechanism to avoid overfitting. While training, the
effectiveness of the classifier was evaluated after every iteration using the
accuracy or F-positive measure. During the training process the following
procedure was followed:

@@

        \item \textbf{Folding} - The training and testing process are run
            multiple times with sections used for various roles. The roles are
            permuted so that the classifier is trained and tested with the
            sections in all possible configurations. In each fold 60\% of the
            partitions are designated as part of the \textit{training set},
            20\% as the \textit{validation set}, and 20\% as the \textit{test set}.
            Since the Naive Bayes classifier does not use a \textit{validation set},
            40\% of the partitions are designated as part of the
            \textit{test set}

@@

The baseline for this project is established by running the Maximum Entropy
classifier without any stopword removal, uncommon word removal (words with a
document frequency $<$ 1), Or feature normalization. The F-score that is
achieved is \textbf{50.72}.

@@

\subsubsection{Number of Features}
\label{ssection:numfeatures}

The number of features retained after feature selection varies from Tweet to Tweet.
To summarize the number of features retained after different processes, Table
\ref{table:numfeatures} presents the mean number of features per Tweet for
positive, neutral, negative examples as well as an overall average.

@@

Interestingly `negative' Tweets contain more features then `positive' Tweets
which, in turn, contain more features than `neutral' tweets. Normalization does
not appear to significantly reduce the number of features per Tweet. This makes
sense since the number of features in the \textit{bag-of-words} would only be
reduced if two or more of the features are normalized to the same thing (e.g.
A Tweet that contained `2' and `3' before normalization would only contain the
feature `\#NUM' after normalization).

@@

As Table \ref{table:maxent-results} shows, the `All' experiment provides the
best results over the baseline. The `Just Stopwords' experiment actually
provides slightly worse results then the baseline indicating that in context of
Tweets with so few features in each example every feature counts. The `Just
Uncommon' experiment does not provide very significant gains when used alone.
This is potentially due to the fact that without normalization turned,
features such as `just' are `Just' are determined to be different and may be
removed if each only occur once.

@@

% TODO make serif?

@@

The implementation created for this project has the ability to show which
features are the most discriminatory. The following is a list of the top 30
most informative features from one run of the Maximum Entropy classifier (note
that some of the weights are negative):

@@

Some of the features listed above make perfect sense, such as `excited' not
being associated with the neutral class and `embarrassing' being associated
with the negative class. However some of the most discriminating features
appear in the dataset infrequently, such as `calvin' and `oneil'.

@@

Table \ref{table:comparison} shows the performance of the classifier produced
for this project compared to results of the workshop participants. The full
table (with out this project's results) can be found in Nakov et al. as Table 9
\cite{Nakov2013}.

@@

As shown in Table \ref{table:comparison}, there is a lot of room for improvement
before this system is a completive solution. However, only simple methods have
been employed including: stopword removal of only 28 words, words with a
document frequency lower than one removed, and normalization of words, emoticons,
and OOV words. These simple preprocessing and feature selection methods in
combination with a Maximum Entropy classifier have demonstrated reasonable
results.

@@

There is room for a number of improvement in the implementation of the system.
The following sections are a number of improvements as future work. Many of the
ideas are from the other submissions to the SemEval-2013 workshop.

@@

The words in the previous Tweet carry no inherent sentiment, however the
message is clearly negative. While a simple approach could be to strip the
octothorpe from the front of the Hashtag (e.g. `\#good' $\rightarrow$ `good'),
the semantic meaning of the two are different. Hashtags often have message
level meaning while simple words have local context withing the sentence. Since
the dataset created for the SemEval-2013 workshop mostly contained examples of
Hashtags used as proper nouns, more annotated Tweets would be needed for
training and testing.

Hashtags sometimes use CamelCase to create multi-word tags (e.g.
`\#SorryNotSorry'). It could prove worthwhile to attempt the decompose
CamelCase Hashtags into its component words during feature selection.

The top solution from the National Research Council of Canada made use of a
Hashtag lexicon that they developed between April of 2012 and December 2012
\cite{Mohammad2013}. The final \textit{NRC Hashtag Sentiment Lexicon} has more
than 350,00 entries and is available at Saif Mohammad's website
\footnote{http://www.saifmohammad.com/WebDocs/NRC-Hashtag-Sentiment-Lexicon-v0.1.zip}.

@@

A challenge that this method could introduce is the introduction of a temporal
element since people's outlooks, and therefore sentiment's in Tweets, will
change over time. As well as public perception of a individual changes, the
prevailing sentiment may change as well.

@@

One problem with the current implementation is that it lack support for
negation. In it's current form the classifier would likely misunderstand the
sequence ``not perfect'' since `not' is a vaguely negative word and `perfect'
is a very positive word. It would likely be classified as positive when in fact
the sentiment is quite clearly negative. A common approach to this problem is
to add a suffix (such as "\_NEG") to word that appear between a negation word
and a section delimiting punctuation mark (such as `,', `.', `:', `;', `!',
`?'). By separating negated features from their non-negated instances, features
can do a better job of distinguishing classes.

@@

One way that many of the other \textit{bag-of-words} solutions capture word order is by
creating features consisting of sequences of words called bigrams. Adding
bigram features adds a large number of features which can add complexity to
training. This concept can also be extended to sequences of length n, called
ngrams.

@@

Most of the more complex solutions submitted to the SemEval-2013 workshop made
uses of sentiment lexicons that capture a term's polarity. These lexicons, such
as SentiWordNet \footnote{http://sentiwordnet.isti.cnr.it/}, express the
positive, neutral, and negative sentiment of each homonym. An example of this
from SentiWordNet, is `good' meaning `morally admirable' having an entirely
positive sentiment versus `good' meaning `having the normally expected amount'
having an entirely neutral sentiment
\footnote{http://sentiwordnet.isti.cnr.it/search.php?q=good}.

@@

Many of the top solutions submitted to the workshop, such as NRC-Canada made
use of a Support Vector Machine for classification \cite{Mohammad2013}.  This
indicates that when used effectively a SVM solution can out-perform a Maximum
Entropy solution. The KLUE team, on the other hand, found that their Maximum
Entropy solution outperform their linear SVM solution \cite{Proisl2013}.

@@

The list of stopwords was acquired online\footnote{www.ranks.nl/stopwords} and is
as follows:

\begin{Verbatim}[fontsize=\footnotesize]
        a
        about
        an
        are
        as
        at
        be
        by
        com
        for
        from
        how
        in
        is
        it
        of
        on
        or
        that
        the
        this
        to
        was
        what
        when
        where
        who
        will
        with
\end{Verbatim}
